PANCHANGE_analysis

### 2020-08-26 PANCHANGE_analysis.Rmd
Gender and Ethnicity collapsed

## Vector
```{r vector with psychiatric disorders}
disorder_columns <- c(
    "depressive_disorders_numeric",
    "anxiety_disorders_numeric",
    "eating_disorders_numeric",
    "obsessive_compulsive_disorders_numeric",
    "psychotic_disorders_numeric",
    "mhd.mania_hypomania_bipolar_or_manicdepression_numeric",
    "mhd.posttraumatic_stress_disorder_ptsd_numeric",
    "autism_spectrum_disorder_numeric",
    "mhd.attention_deficit_hyperactivity_disorder_numeric",
    "mhd.personality_disorder_numeric",
    "control_numeric"
)
```

@@

```{r gather to long format Disorder_hierarchical}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_inc_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    all_of(diff_score_cols),
    Disorder_inc_raw,
    Diagnosis
  ) %>%
  mutate(Disorder_inc =
           recode_factor(
             Disorder_inc_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             autism_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

@@

Ethnicity should be recoded
European
Mixed or multiple ethnic origins
African or African British
Asian or Asian British
Arab
```{r Disorder_inc by Ethnicity}
Disorder_inc_by_Ethnicity <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Ethnicity) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r Disorder_inc by Ethnicity plot}
Disorder_inc_by_Ethnicity %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
  ) +
  geom_bar(
    aes(fill = Ethnicity),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Ethnicity") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6,
    #    labels = c("Prepandemic", "Retrospective")
  ) +
  coord_flip()
```

@@

```{r gad gather to long format Time_points gad.diff_score}
dat.long.gad.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "gad.diff_score",
    all_of(gad.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    gad.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "gad.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r gad gather to long format Time_points gad.sum_score}
dat.long.gad.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "gad.sum_score",
    all_of(gad.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    gad.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "gad.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r phq gather to long format Time_points}
dat.long.phq.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.diff_score",
    all_of(phq.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    phq.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r phq gather to long format Time_points phq.sum_score}
dat.long.phq.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.sum_score",
    all_of(phq.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    phq.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r ocir gather to long format Time_points}
dat.long.ocir.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "ocir.diff_score",
    all_of(ocir.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    ocir.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "ocir.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r ocir gather to long format Time_points ocir.sum_score}
dat.long.ocir.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "ocir.sum_score",
    all_of(ocir.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    ocir.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "ocir.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r pcl gather to long format Time_points pcl.sum_score}
dat.long.pcl.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "pcl.sum_score",
    all_of(pcl.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Disorder_hierarchical,
    Time_point_raw,
    pcl.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "pcl.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

Model 1: Model 0 + Sex + Age + Ethnicity + Time registration and baseline
```{r}
model1 <- lm(
  formula = phq.diff_score_base_prepan ~
    anxiety_disorders +
    depressive_disorders +
    eating_disorders +
    obsessive_compulsive_disorders +
    psychotic_disorders +
    mhd.mania_hypomania_bipolar_or_manicdepression +
    mhd.posttraumatic_stress_disorder_ptsd +
    autism_spectrum_disorder +
    mhd.attention_deficit_hyperactivity_disorder +
    mhd.personality_disorder +
    Gender + age_category + Ethnicity + time_diff_sign_up_coping,
  data = dat)

@@

Model 1: Model 0 + Sex + Age + Ethnicity + Time registration and baseline
```{r}
model1 <- lm(
  formula = pcl.diff_score_base_prepan ~ Disorder_hierarchical + Gender + age_category + Ethnicity + time_diff_sign_up_coping,
  data = dat)

@@

Model 1: Model 0 + Sex + Age + Ethnicity + Time registration and baseline
```{r}
model1 <- lm(
  formula = gad.diff_score_base_prepan ~ Disorder + Gender + Age + Ethnicity + time_diff_sign_up_coping,
  data = dat)

>>>>

## Vector
```{r vector with psychiatric disorders}
disorder_columns <- c(
  "control_numeric",  
  "depressive_disorders_numeric",
    "anxiety_disorders_numeric",
    "mhd.posttraumatic_stress_disorder_ptsd_numeric",
    "eating_disorders_numeric",
    "obsessive_compulsive_disorders_numeric",
    "mhd.attention_deficit_hyperactivity_disorder_numeric",
    "mhd.personality_disorder_numeric",
    "autism_spectrum_disorder_numeric",
    "mhd.mania_hypomania_bipolar_or_manicdepression_numeric",
    "psychotic_disorders_numeric"
)
```

@@

```{r gather to long format Disorder_hierarchical}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_inc_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    all_of(diff_score_cols),
    Disorder_inc_raw,
    Diagnosis
  ) %>%
  mutate(Disorder_inc =
           recode_factor(
             Disorder_inc_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             autism_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

@@

Ethnicity should be recoded
European
Mixed or multiple ethnic origins
African or African British
Asian or Asian British
Arab
```{r Disorder_inc by Ethnicity_collapsed}
Disorder_inc_by_Ethnicity <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  count(Disorder_inc, Ethnicity_collapsed) %>%
  mutate(Prop = round(n/sum(n), 2))

@@

```{r Disorder_inc by Ethnicity_collapsed plot}
Disorder_inc_by_Ethnicity %>%
  ggplot(mapping = aes(
    x = Disorder_inc,
    y = n)
  ) +
  geom_bar(
    aes(fill = Ethnicity_collapsed),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorders",
    y = "Number of participants",
    fill = "Ethnicity_collapsed") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6,
    #    labels = c("Prepandemic", "Retrospective")
  ) +
  coord_flip()
```

@@

```{r gad gather to long format Time_points gad.diff_score}
dat.long.gad.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "gad.diff_score",
    all_of(gad.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    Disorder_hierarchical,
    Time_point_raw,
    gad.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "gad.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r gad gather to long format Time_points gad.sum_score}
dat.long.gad.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "gad.sum_score",
    all_of(gad.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    Disorder_hierarchical,
    Time_point_raw,
    gad.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "gad.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r phq gather to long format Time_points}
dat.long.phq.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.diff_score",
    all_of(phq.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    Disorder_hierarchical,
    Time_point_raw,
    phq.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r phq gather to long format Time_points phq.sum_score}
dat.long.phq.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.sum_score",
    all_of(phq.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    Disorder_hierarchical,
    Time_point_raw,
    phq.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r ocir gather to long format Time_points}
dat.long.ocir.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "ocir.diff_score",
    all_of(ocir.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    Disorder_hierarchical,
    Time_point_raw,
    ocir.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "ocir.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r ocir gather to long format Time_points ocir.sum_score}
dat.long.ocir.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "ocir.sum_score",
    all_of(ocir.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    Disorder_hierarchical,
    Time_point_raw,
    ocir.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "ocir.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

```{r pcl gather to long format Time_points pcl.sum_score}
dat.long.pcl.sum <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "pcl.sum_score",
    all_of(pcl.sum_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender_collapsed,
    Ethnicity_collapsed,
    Disorder_hierarchical,
    Time_point_raw,
    pcl.sum_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "pcl.sum_score_", "")) %>%
  select(-Time_point_raw)

@@

Model 1: Model 0 + Sex + Age + Ethnicity_collapsed + Time registration and baseline
```{r}
model1 <- lm(
  formula = phq.diff_score_base_prepan ~
    anxiety_disorders +
    depressive_disorders +
    eating_disorders +
    obsessive_compulsive_disorders +
    psychotic_disorders +
    mhd.mania_hypomania_bipolar_or_manicdepression +
    mhd.posttraumatic_stress_disorder_ptsd +
    autism_spectrum_disorder +
    mhd.attention_deficit_hyperactivity_disorder +
    mhd.personality_disorder +
    Gender_collapsed + age_category + Ethnicity_collapsed + time_diff_sign_up_coping,
  data = dat)

@@

Model 1: Model 0 + Sex + Age + Ethnicity_collapsed + Time registration and baseline
```{r}
model1 <- lm(
  formula = pcl.diff_score_base_prepan ~ Disorder_hierarchical + Gender_collapsed + age_category + Ethnicity_collapsed + time_diff_sign_up_coping,
  data = dat)

@@

Model 1: Model 0 + Sex + Age + Ethnicity_collapsed + Time registration and baseline
```{r}
model1 <- lm(
  formula = gad.diff_score_base_prepan ~ Disorder + Gender_collapsed + Age + Ethnicity_collapsed + time_diff_sign_up_coping,
  data = dat)
